New cardinality estimation algorithms for HyperLogLog sketches

### 2016-12-20 paper/paper.tex
added task for generation of arXiv upload package, made cleveref options compatible with arXiv processing
minor changes

\documentclass[a4paper]{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{algpseudocode,algorithm,algorithmicx}
\usepackage{float}
\usepackage{url}
\usepackage{booktabs}
\usepackage{array}
\usepackage{paralist}
\usepackage{verbatim}
\usepackage{subfig}
\usepackage{hyperref}
\usepackage[binary-units = true]{siunitx}
\usepackage{csvsimple}
\usepackage{multirow}
\usepackage[capitalise,nameinlink,noabbrev]{cleveref}

@@

\subsection{Maximum likelihood estimation algorithm}
In order to get a fast cardinality estimation algorithm, it is crucial to minimize the evaluation costs for \eqref{equ:funcprime}. A couple of optimizations allow significant reduction of the computational effort:
\begin{itemize}
\item Only a fraction of all count values $\symCountVariate_\symRegVal$ is non-zero. If we denote $\symRegValVariate_\text{min}:=\min\lbrace \symRegVal\vert\symCountVariate_\symRegVal>0\rbrace$ and $\symRegValVariate_\text{max}:=\max\lbrace \symRegVal\vert\symCountVariate_\symRegVal>0\rbrace$,  it is sufficient to loop over all indices in the range $[\symRegValVariate_\text{min}, \symRegValVariate_\text{max}]$.
\item The sum $\sum_{\symRegVal=0}^\symRegRange \frac{\symCountVariate_\symRegVal}{2^\symRegVal}$ in \eqref{equ:funcprime} can be precalculated and reused for all function evaluations.
\item Many programming languages allow the efficient multiplication and division by any integral power of two using special functions, such as \texttt{ldexp} in C/C++ or \texttt{scalb} in Java.
\item The function $\symHelper(\symX)$ only needs to be evaluated at values $\left\lbrace\frac{\symX}{2^{\symRegValVariate'_\text{max}}},\frac{\symX}{2^{\symRegValVariate'_\text{max}-1}},\ldots,\frac{\symX}{2^{\symRegValVariate_\text{min}}}\right\rbrace$ where $\symRegValVariate'_\text{max} := \min(\symRegValVariate_\text{max}, \symRegRange)$. This series corresponds to a geometric series with ratio two. A straightforward calculation using \eqref{equ:helper} is very expensive because of the exponential function. However, if we know $\symHelper\!\left(\frac{\symX}{2^{\symRegValVariate'_\text{max}}}\right)$ all other required function values can be easily obtained using the identity
\begin{equation}
\label{equ:helper_recursion1}
\symHelper(4\symX) = \frac{\symX+\symHelper(2\symX)\left(1-\symHelper(2\symX)\right)}{\symX+\left(1-\symHelper(2\symX)\right)}.
\end{equation}
%or
%\begin{equation}
%\label{equ:helper_recursion2}
%\symHelper\!\left(\frac{\symX}{2^{\symRegVal}}\right) = \frac{\frac{\symX}{2^{\symRegVal+2}}+\symHelper\!\left(\frac{\symX}{2^{\symRegVal+1}}\right)\left(1-\symHelper\!\left(\frac{\symX}{2^{\symRegVal+1}}\right)\right)}{\frac{\symX}{2^{\symRegVal+2}}+\left(1-\symHelper\!\left(\frac{\symX}{2^{\symRegVal+1}}\right)\right)}
%\end{equation}
This recursive formula is stable in a sense that the relative error of $\symHelper(4\symX)$ is smaller than that of $\symHelper(2\symX)$ as shown in \cref{app:helper_stable}.

@@

\begin{table}
\centering
\caption{List of true cardinalities of the pairwise disjoint sets $\symSetA$, $\symSetB$, and $\symSetX$, for which maximum likelihood estimation from two HyperLogLog sketches with parameters $\symPrecision=20$ and $\symRegRange=44$ representing sets $\symSetS_1 = \symSetA\cup\symSetX$ and  $\symSetS_2 = \symSetB\cup\symSetX$ was investigated. The last column shows the average number of iteration cycles until the stop criterion of the BFGS algorithm was satisfied.
}
\label{tbl:joint_estimation_cases}
\csvreader[before reading=\scriptsize, tabular=r|rrr|rr|rr, table head= & 
\multicolumn{3}{c|}{true cardinalities} &
\multicolumn{1}{c}{Jaccard index} &
\multicolumn{1}{c|}{cardinality ratio} &
\multicolumn{1}{c}{avg number}
\\
\# &  
\multicolumn{1}{c}{$|\symSetA|$} & 
\multicolumn{1}{c}{$|\symSetB|$} & 
\multicolumn{1}{c|}{$|\symSetX|$} & 
\multicolumn{1}{c}{$|\symSetS_1\cap\symSetS_2|/|\symSetS_1\cup\symSetS_2|$} & 
\multicolumn{1}{c|}{$|\symSetS_1|/|\symSetS_2|$} &
\multicolumn{1}{c}{iterations}
\\\hline]
{joint_cardinality_calculation.csv}{
trueCardA=\trueCardA,
trueCardB=\trueCardB,
trueCardX=\trueCardX,
trueJaccardIdx=\trueJaccardIdx,
trueRatio=\trueRatio,
avgNumEvaluations=\avgNumEvaluations
}{
\thecsvrow & 
\num{\trueCardA} & 
\num{\trueCardB} & 
\num{\trueCardX} & 
\numformat{\trueJaccardIdx} & 
\numformattwo{\trueRatio} &
\numformattwo{\avgNumEvaluations}
}
\end{table}

\begin{table}
\centering
\caption{The mean, the standard deviation, and the root-mean-square error of the relative estimation error are given for all the cases given in \cref{tbl:joint_estimation_cases} when estimating $|\symSetA|$ using the conventional approach and the maximum likelihood method, respectively. The root-mean-square error improvement factor of the maximum likelihood method is shown in the last column.}
\label{tbl:joint_estimation_a}
\csvreader[before reading=\scriptsize, tabular=r|rrr|rrr|r, table head=
 & 
\multicolumn{3}{c|}{conventional approach}
&
\multicolumn{3}{c|}{maximum likelihood method}
&
\multicolumn{1}{c}{improvement}
\\
\# & 
\multicolumn{1}{c}{mean} & 
\multicolumn{1}{c}{stdev} & 
\multicolumn{1}{c|}{rmse} & 
\multicolumn{1}{c}{mean} & 
\multicolumn{1}{c}{stdev} & 
\multicolumn{1}{c|}{rmse} &
\multicolumn{1}{c}{rmse ratio}
\\\hline]
{joint_cardinality_calculation.csv}{
inclExclMeanA=\inclExclMeanA,
inclExclStdDevA=\inclExclStdDevA,
inclExclRMSEA=\inclExclRMSEA,
maxLikeMeanA=\maxLikeMeanA,
maxLikeStdDevA=\maxLikeStdDevA,
maxLikeRMSEA=\maxLikeRMSEA,
improvementRmseA=\improvementRmseA
}{
\thecsvrow & 
\numformat{\inclExclMeanA} & 
\numformat{\inclExclStdDevA} & 
\numformat{\inclExclRMSEA} & 
\numformat{\maxLikeMeanA} & 
\numformat{\maxLikeStdDevA} & 
\numformat{\maxLikeRMSEA} &
\numformattwo{\improvementRmseA}
}
\end{table}

\begin{table}
\centering
\caption{The mean, the standard deviation, and the root-mean-square error of the relative estimation error are given for all the cases given in \cref{tbl:joint_estimation_cases} when estimating $|\symSetB|$ using the conventional approach and the maximum likelihood method, respectively. The root-mean-square error improvement factor of the maximum likelihood method is shown in the last column.}
\label{tbl:joint_estimation_b}
\csvreader[before reading=\scriptsize, tabular=r|rrr|rrr|r, table head=
 & 
\multicolumn{3}{c|}{conventional approach}
&
\multicolumn{3}{c|}{maximum likelihood method}
&
\multicolumn{1}{c}{improvement}
\\
\# & 
\multicolumn{1}{c}{mean} & 
\multicolumn{1}{c}{stdev} & 
\multicolumn{1}{c|}{rmse} & 
\multicolumn{1}{c}{mean} & 
\multicolumn{1}{c}{stdev} & 
\multicolumn{1}{c|}{rmse} &
\multicolumn{1}{c}{rmse ratio}
\\\hline]
{joint_cardinality_calculation.csv}{
inclExclMeanB=\inclExclMeanB,
inclExclStdDevB=\inclExclStdDevB,
inclExclRMSEB=\inclExclRMSEB,
maxLikeMeanB=\maxLikeMeanB,
maxLikeStdDevB=\maxLikeStdDevB,
maxLikeRMSEB=\maxLikeRMSEB,
improvementRmseB=\improvementRmseB
}{
\thecsvrow & 
\numformat{\inclExclMeanB} & 
\numformat{\inclExclStdDevB} & 
\numformat{\inclExclRMSEB} & 
\numformat{\maxLikeMeanB} & 
\numformat{\maxLikeStdDevB} & 
\numformat{\maxLikeRMSEB} &
\numformattwo{\improvementRmseB}
}
\end{table}

\begin{table}
\centering
\caption{The mean, the standard deviation, and the root-mean-square error of the relative estimation error are given for all the cases given in \cref{tbl:joint_estimation_cases} when estimating $|\symSetX|$ using the conventional approach and the maximum likelihood method, respectively. The root-mean-square error improvement factor of the maximum likelihood method is shown in the last column.}
\label{tbl:joint_estimation_x}
\csvreader[before reading=\scriptsize, tabular=r|rrr|rrr|r, table head=
 & 
\multicolumn{3}{c|}{conventional approach}
&
\multicolumn{3}{c|}{maximum likelihood method}
&
\multicolumn{1}{c}{improvement}
\\
\# & 
\multicolumn{1}{c}{mean} & 
\multicolumn{1}{c}{stdev} & 
\multicolumn{1}{c|}{rmse} & 
\multicolumn{1}{c}{mean} & 
\multicolumn{1}{c}{stdev} & 
\multicolumn{1}{c|}{rmse} &
\multicolumn{1}{c}{rmse ratio}
\\\hline]
{joint_cardinality_calculation.csv}{
inclExclMeanX=\inclExclMeanX,
inclExclStdDevX=\inclExclStdDevX,
inclExclRMSEX=\inclExclRMSEX,
maxLikeMeanX=\maxLikeMeanX,
maxLikeStdDevX=\maxLikeStdDevX,
maxLikeRMSEX=\maxLikeRMSEX,
improvementRmseX=\improvementRmseX
}{
\thecsvrow & 
\numformat{\inclExclMeanX} & 
\numformat{\inclExclStdDevX} & 
\numformat{\inclExclRMSEX} & 
\numformat{\maxLikeMeanX} & 
\numformat{\maxLikeStdDevX} & 
\numformat{\maxLikeRMSEX} &
\numformattwo{\improvementRmseX}
}
\end{table}

\begin{table}
\centering
\caption{The mean, the standard deviation, and the root-mean-square error of the relative estimation error are given for all the cases given in \cref{tbl:joint_estimation_cases} when estimating $|\symSetA\cup\symSetB\cup\symSetX|$ using the conventional approach and the maximum likelihood method, respectively. The root-mean-square error improvement factor of the maximum likelihood method is shown in the last column.}
\label{tbl:joint_estimation_abx}
\csvreader[before reading=\scriptsize, tabular=r|rrr|rrr|r, table head=
 & 
\multicolumn{3}{c|}{conventional approach}
&
\multicolumn{3}{c|}{maximum likelihood method}
&
\multicolumn{1}{c}{improvement}
\\
\# & 
\multicolumn{1}{c}{mean} & 
\multicolumn{1}{c}{stdev} & 
\multicolumn{1}{c|}{rmse} & 
\multicolumn{1}{c}{mean} & 
\multicolumn{1}{c}{stdev} & 
\multicolumn{1}{c|}{rmse} &
\multicolumn{1}{c}{rmse ratio}
\\\hline]
{joint_cardinality_calculation.csv}{
inclExclMeanABX=\inclExclMeanABX,
inclExclStdDevABX=\inclExclStdDevABX,
inclExclRMSEABX=\inclExclRMSEABX,
maxLikeMeanABX=\maxLikeMeanABX,
maxLikeStdDevABX=\maxLikeStdDevABX,
maxLikeRMSEABX=\maxLikeRMSEABX,
improvementRmseABX=\improvementRmseABX
}{
\thecsvrow & 
\numformat{\inclExclMeanABX} & 
\numformat{\inclExclStdDevABX} & 
\numformat{\inclExclRMSEABX} & 
\numformat{\maxLikeMeanABX} & 
\numformat{\maxLikeStdDevABX} & 
\numformat{\maxLikeRMSEABX} &
\numformattwo{\improvementRmseABX}
}
\end{table}

%\begin{table}
%\centering
%\caption{TODO}
%\label{tbl:joint_estimation_jaccard}
%\csvreader[before reading=\scriptsize, tabular=r|rrr|rrr|r, table head=
% & 
%\multicolumn{3}{c|}{conventional approach}
%&
%\multicolumn{3}{c|}{maximum likelihood method}
%&
%\multicolumn{1}{c}{improvement}
%\\
%\# & 
%\multicolumn{1}{c}{mean} & 
%\multicolumn{1}{c}{stdev} & 
%\multicolumn{1}{c|}{rmse} & 
%\multicolumn{1}{c}{mean} & 
%\multicolumn{1}{c}{stdev} & 
%\multicolumn{1}{c|}{rmse} &
%\multicolumn{1}{c}{rmse ratio}
%\\\hline]
%{joint_cardinality_calculation.csv}{
%inclExclMeanJaccardIdx=\inclExclMeanJaccardIdx,
%inclExclStdDevJaccardIdx=\inclExclStdDevJaccardIdx,
%inclExclRMSEJaccardIdx=\inclExclRMSEJaccardIdx,
%maxLikeMeanJaccardIdx=\maxLikeMeanJaccardIdx,
%maxLikeStdDevJaccardIdx=\maxLikeStdDevJaccardIdx,
%maxLikeRMSEJaccardIdx=\maxLikeRMSEJaccardIdx,
%improvementRmseJaccard=\improvementRmseJaccard
%}{
%\thecsvrow & 
%\numformat{\inclExclMeanJaccardIdx} & 
%\numformat{\inclExclStdDevJaccardIdx} & 
%\numformat{\inclExclRMSEJaccardIdx} & 
%\numformat{\maxLikeMeanJaccardIdx} & 
%\numformat{\maxLikeStdDevJaccardIdx} & 
%\numformat{\maxLikeRMSEJaccardIdx} &
%\numformattwo{\improvementRmseJaccard}
%}
%\end{table}

@@

\section{Analysis of $\symPowerSeriesFunc(\symX)$}
\label{app:analysis_xi}
The Fourier series of the function $\symPowerSeriesFunc(\symX)$, which is defined by
\begin{equation}
\label{equ:def_xi}
\symPowerSeriesFunc(\symX):= \log(2) \sum_{\symRegVal = -\infty}^\infty
2^{\symRegVal+\symX}
e^{-2^{\symRegVal+\symX}}
\end{equation}
and which has period equal to 1, is
\begin{equation}
\symPowerSeriesFunc(\symX) = \frac{\symA_0}{2} + \Re\!\left(\sum_{\symIndexL=1}^{\infty}
\symA_\symIndexL e^{2\pi\symImaginary\symIndexL\symX}\right)
\end{equation}
with coefficients
\begin{multline}
\symA_\symIndexL = 2\int_0^1 \symPowerSeriesFunc(\symX) e^{-2\pi\symImaginary\symIndexL\symX}d\symX
=
2\log(2)
\int_0^1
\sum_{\symRegVal = -\infty}^\infty
2^{\symRegVal+\symX}
e^{-2^{\symRegVal+\symX}}
e^{-2\pi\symImaginary\symIndexL\symX}
d\symX
=
\\
=
2\log(2)
\sum_{\symRegVal = -\infty}^\infty
\int_{\symRegVal}^{\symRegVal+1}
2^{\symX}
e^{-2^{\symX}}
e^{-2\pi\symImaginary\symIndexL\symX}
d\symX
=
2\log(2) \int_{-\infty}^{\infty} 2^\symX
e^{-2^\symX}
e^{-2\pi\symImaginary\symIndexL\symX}d\symX.
\end{multline}
The variable transformation $\symY = 2^\symX$ yields
\begin{equation}
\symA_\symIndexL
=
2\int_{0}^{\infty} e^{-\symY}
\symY^{-\frac{2\pi\symImaginary\symIndexL}{\log(2)}}d\symY
= 
2\Gamma\!\left(1-\frac{2\pi\symImaginary\symIndexL}{\log(2)}\right)
\end{equation}
where $\Gamma$ denotes the gamma function. Obviously, the constant term in the Fourier series is equal to $1$, because $\symA_0=2$.

>>>>

\documentclass[a4paper]{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{algpseudocode,algorithm,algorithmicx}
\usepackage{float}
\usepackage{url}
\usepackage{booktabs}
\usepackage{array}
\usepackage{paralist}
\usepackage{verbatim}
\usepackage{subfig}
\usepackage{hyperref}
\usepackage[binary-units = true]{siunitx}
\usepackage{csvsimple}
\usepackage{multirow}
\usepackage[capitalise]{cleveref}

@@

\subsection{Maximum likelihood estimation algorithm}
In order to get a fast cardinality estimation algorithm, it is crucial to minimize evaluation costs for \eqref{equ:funcprime}. A couple of optimizations allow significant reduction of computational effort:
\begin{itemize}
\item Only a fraction of all count values $\symCountVariate_\symRegVal$ is non-zero. If we denote $\symRegValVariate_\text{min}:=\min\lbrace \symRegVal\vert\symCountVariate_\symRegVal>0\rbrace$ and $\symRegValVariate_\text{max}:=\max\lbrace \symRegVal\vert\symCountVariate_\symRegVal>0\rbrace$,  it is sufficient to loop over all indices in the range $[\symRegValVariate_\text{min}, \symRegValVariate_\text{max}]$.
\item The sum $\sum_{\symRegVal=0}^\symRegRange \frac{\symCountVariate_\symRegVal}{2^\symRegVal}$ in \eqref{equ:funcprime} can be precalculated and reused for all function evaluations.
\item Many programming languages allow the efficient multiplication and division by any integral power of two using special functions, such as \texttt{ldexp} in C/C++ or \texttt{scalb} in Java.
\item The function $\symHelper(\symX)$ only needs to be evaluated at values $\left\lbrace\frac{\symX}{2^{\symRegValVariate'_\text{max}}},\frac{\symX}{2^{\symRegValVariate'_\text{max}-1}},\ldots,\frac{\symX}{2^{\symRegValVariate_\text{min}}}\right\rbrace$ where $\symRegValVariate'_\text{max} := \min(\symRegValVariate_\text{max}, \symRegRange)$. This series corresponds to a geometric series with ratio two. A straightforward calculation using \eqref{equ:helper} is very expensive because of the exponential function. However, if we know $\symHelper\!\left(\frac{\symX}{2^{\symRegValVariate'_\text{max}}}\right)$ all other required function values can be easily obtained using the identity
\begin{equation}
\label{equ:helper_recursion1}
\symHelper(4\symX) = \frac{\symX+\symHelper(2\symX)\left(1-\symHelper(2\symX)\right)}{\symX+\left(1-\symHelper(2\symX)\right)}.
\end{equation}
%or
%\begin{equation}
%\label{equ:helper_recursion2}
%\symHelper\!\left(\frac{\symX}{2^{\symRegVal}}\right) = \frac{\frac{\symX}{2^{\symRegVal+2}}+\symHelper\!\left(\frac{\symX}{2^{\symRegVal+1}}\right)\left(1-\symHelper\!\left(\frac{\symX}{2^{\symRegVal+1}}\right)\right)}{\frac{\symX}{2^{\symRegVal+2}}+\left(1-\symHelper\!\left(\frac{\symX}{2^{\symRegVal+1}}\right)\right)}
%\end{equation}
This recursive formula is stable in a sense that the relative error of $\symHelper(4\symX)$ is smaller than that of $\symHelper(2\symX)$ as shown in \cref{app:helper_stable}.

@@

\begin{table}
\centering
\captionabove{List of true cardinalities of the pairwise disjoint sets $\symSetA$, $\symSetB$, and $\symSetX$, for which maximum likelihood estimation from two HyperLogLog sketches with parameters $\symPrecision=20$ and $\symRegRange=44$ representing sets $\symSetS_1 = \symSetA\cup\symSetX$ and  $\symSetS_2 = \symSetB\cup\symSetX$ was investigated. The last column shows the average number of iteration cycles until the stop criterion of the BFGS algorithm was satisfied.
}
\label{tbl:joint_estimation_cases}
\csvreader[before reading=\scriptsize, tabular=r|rrr|rr|rr, table head= & 
\multicolumn{3}{c|}{true cardinalities} &
\multicolumn{1}{c}{Jaccard index} &
\multicolumn{1}{c|}{cardinality ratio} &
\multicolumn{1}{c}{avg number}
\\
\# &  
\multicolumn{1}{c}{$|\symSetA|$} & 
\multicolumn{1}{c}{$|\symSetB|$} & 
\multicolumn{1}{c|}{$|\symSetX|$} & 
\multicolumn{1}{c}{$|\symSetS_1\cap\symSetS_2|/|\symSetS_1\cup\symSetS_2|$} & 
\multicolumn{1}{c|}{$|\symSetS_1|/|\symSetS_2|$} &
\multicolumn{1}{c}{iterations}
\\\hline]
{joint_cardinality_calculation.csv}{
trueCardA=\trueCardA,
trueCardB=\trueCardB,
trueCardX=\trueCardX,
trueJaccardIdx=\trueJaccardIdx,
trueRatio=\trueRatio,
avgNumEvaluations=\avgNumEvaluations
}{
\thecsvrow & 
\num{\trueCardA} & 
\num{\trueCardB} & 
\num{\trueCardX} & 
\numformat{\trueJaccardIdx} & 
\numformattwo{\trueRatio} &
\numformattwo{\avgNumEvaluations}
}
\end{table}

\begin{table}
\centering
\captionabove{The mean, the standard deviation, and the root-mean-square error of the relative estimation error are given for all the cases given in \cref{tbl:joint_estimation_cases} when estimating $|\symSetA|$ using the conventional approach and the maximum likelihood method, respectively. The root-mean-square error improvement factor of the maximum likelihood method is shown in the last column.}
\label{tbl:joint_estimation_a}
\csvreader[before reading=\scriptsize, tabular=r|rrr|rrr|r, table head=
 & 
\multicolumn{3}{c|}{conventional approach}
&
\multicolumn{3}{c|}{maximum likelihood method}
&
\multicolumn{1}{c}{improvement}
\\
\# & 
\multicolumn{1}{c}{mean} & 
\multicolumn{1}{c}{stdev} & 
\multicolumn{1}{c|}{rmse} & 
\multicolumn{1}{c}{mean} & 
\multicolumn{1}{c}{stdev} & 
\multicolumn{1}{c|}{rmse} &
\multicolumn{1}{c}{rmse ratio}
\\\hline]
{joint_cardinality_calculation.csv}{
inclExclMeanA=\inclExclMeanA,
inclExclStdDevA=\inclExclStdDevA,
inclExclRMSEA=\inclExclRMSEA,
maxLikeMeanA=\maxLikeMeanA,
maxLikeStdDevA=\maxLikeStdDevA,
maxLikeRMSEA=\maxLikeRMSEA,
improvementRmseA=\improvementRmseA
}{
\thecsvrow & 
\numformat{\inclExclMeanA} & 
\numformat{\inclExclStdDevA} & 
\numformat{\inclExclRMSEA} & 
\numformat{\maxLikeMeanA} & 
\numformat{\maxLikeStdDevA} & 
\numformat{\maxLikeRMSEA} &
\numformattwo{\improvementRmseA}
}
\end{table}

\begin{table}
\centering
\captionabove{The mean, the standard deviation, and the root-mean-square error of the relative estimation error are given for all the cases given in \cref{tbl:joint_estimation_cases} when estimating $|\symSetB|$ using the conventional approach and the maximum likelihood method, respectively. The root-mean-square error improvement factor of the maximum likelihood method is shown in the last column.}
\label{tbl:joint_estimation_b}
\csvreader[before reading=\scriptsize, tabular=r|rrr|rrr|r, table head=
 & 
\multicolumn{3}{c|}{conventional approach}
&
\multicolumn{3}{c|}{maximum likelihood method}
&
\multicolumn{1}{c}{improvement}
\\
\# & 
\multicolumn{1}{c}{mean} & 
\multicolumn{1}{c}{stdev} & 
\multicolumn{1}{c|}{rmse} & 
\multicolumn{1}{c}{mean} & 
\multicolumn{1}{c}{stdev} & 
\multicolumn{1}{c|}{rmse} &
\multicolumn{1}{c}{rmse ratio}
\\\hline]
{joint_cardinality_calculation.csv}{
inclExclMeanB=\inclExclMeanB,
inclExclStdDevB=\inclExclStdDevB,
inclExclRMSEB=\inclExclRMSEB,
maxLikeMeanB=\maxLikeMeanB,
maxLikeStdDevB=\maxLikeStdDevB,
maxLikeRMSEB=\maxLikeRMSEB,
improvementRmseB=\improvementRmseB
}{
\thecsvrow & 
\numformat{\inclExclMeanB} & 
\numformat{\inclExclStdDevB} & 
\numformat{\inclExclRMSEB} & 
\numformat{\maxLikeMeanB} & 
\numformat{\maxLikeStdDevB} & 
\numformat{\maxLikeRMSEB} &
\numformattwo{\improvementRmseB}
}
\end{table}

\begin{table}
\centering
\captionabove{The mean, the standard deviation, and the root-mean-square error of the relative estimation error are given for all the cases given in \cref{tbl:joint_estimation_cases} when estimating $|\symSetX|$ using the conventional approach and the maximum likelihood method, respectively. The root-mean-square error improvement factor of the maximum likelihood method is shown in the last column.}
\label{tbl:joint_estimation_x}
\csvreader[before reading=\scriptsize, tabular=r|rrr|rrr|r, table head=
 & 
\multicolumn{3}{c|}{conventional approach}
&
\multicolumn{3}{c|}{maximum likelihood method}
&
\multicolumn{1}{c}{improvement}
\\
\# & 
\multicolumn{1}{c}{mean} & 
\multicolumn{1}{c}{stdev} & 
\multicolumn{1}{c|}{rmse} & 
\multicolumn{1}{c}{mean} & 
\multicolumn{1}{c}{stdev} & 
\multicolumn{1}{c|}{rmse} &
\multicolumn{1}{c}{rmse ratio}
\\\hline]
{joint_cardinality_calculation.csv}{
inclExclMeanX=\inclExclMeanX,
inclExclStdDevX=\inclExclStdDevX,
inclExclRMSEX=\inclExclRMSEX,
maxLikeMeanX=\maxLikeMeanX,
maxLikeStdDevX=\maxLikeStdDevX,
maxLikeRMSEX=\maxLikeRMSEX,
improvementRmseX=\improvementRmseX
}{
\thecsvrow & 
\numformat{\inclExclMeanX} & 
\numformat{\inclExclStdDevX} & 
\numformat{\inclExclRMSEX} & 
\numformat{\maxLikeMeanX} & 
\numformat{\maxLikeStdDevX} & 
\numformat{\maxLikeRMSEX} &
\numformattwo{\improvementRmseX}
}
\end{table}

\begin{table}
\centering
\captionabove{The mean, the standard deviation, and the root-mean-square error of the relative estimation error are given for all the cases given in \cref{tbl:joint_estimation_cases} when estimating $|\symSetA\cup\symSetB\cup\symSetX|$ using the conventional approach and the maximum likelihood method, respectively. The root-mean-square error improvement factor of the maximum likelihood method is shown in the last column.}
\label{tbl:joint_estimation_abx}
\csvreader[before reading=\scriptsize, tabular=r|rrr|rrr|r, table head=
 & 
\multicolumn{3}{c|}{conventional approach}
&
\multicolumn{3}{c|}{maximum likelihood method}
&
\multicolumn{1}{c}{improvement}
\\
\# & 
\multicolumn{1}{c}{mean} & 
\multicolumn{1}{c}{stdev} & 
\multicolumn{1}{c|}{rmse} & 
\multicolumn{1}{c}{mean} & 
\multicolumn{1}{c}{stdev} & 
\multicolumn{1}{c|}{rmse} &
\multicolumn{1}{c}{rmse ratio}
\\\hline]
{joint_cardinality_calculation.csv}{
inclExclMeanABX=\inclExclMeanABX,
inclExclStdDevABX=\inclExclStdDevABX,
inclExclRMSEABX=\inclExclRMSEABX,
maxLikeMeanABX=\maxLikeMeanABX,
maxLikeStdDevABX=\maxLikeStdDevABX,
maxLikeRMSEABX=\maxLikeRMSEABX,
improvementRmseABX=\improvementRmseABX
}{
\thecsvrow & 
\numformat{\inclExclMeanABX} & 
\numformat{\inclExclStdDevABX} & 
\numformat{\inclExclRMSEABX} & 
\numformat{\maxLikeMeanABX} & 
\numformat{\maxLikeStdDevABX} & 
\numformat{\maxLikeRMSEABX} &
\numformattwo{\improvementRmseABX}
}
\end{table}

@@

\section{Analysis of $\symPowerSeriesFunc(\symX)$}
\label{app:analysis_xi}
The Fourier series of the periodic function
\begin{equation}
\label{equ:def_xi}
\symPowerSeriesFunc(\symX):= \log(2) \sum_{\symRegVal = -\infty}^\infty
2^{\symRegVal+\symX}
e^{-2^{\symRegVal+\symX}},
\end{equation}
which has a period equal to 1, is
\begin{equation}
\symPowerSeriesFunc(\symX) = \frac{\symA_0}{2} + \Re\!\left(\sum_{\symIndexL=1}^{\infty}
\symA_\symIndexL e^{2\pi\symImaginary\symIndexL\symX}\right)
\end{equation}
with coefficients
\begin{multline}
\symA_\symIndexL = 2\int_0^1 \symPowerSeriesFunc(\symX) e^{-2\pi\symImaginary\symIndexL\symX}d\symX
=
2\log(2)
\int_0^1
\sum_{\symRegVal = -\infty}^\infty
2^{\symRegVal+\symX}
e^{-2^{\symRegVal+\symX}}
e^{-2\pi\symImaginary\symIndexL\symX}
d\symX
=
\\
=
2\log(2)
\sum_{\symRegVal = -\infty}^\infty
\int_{\symRegVal}^{\symRegVal+1}
2^{\symX}
e^{-2^{\symX}}
e^{-2\pi\symImaginary\symIndexL\symX}
d\symX
=
2\log(2) \int_{-\infty}^{\infty} 2^\symX
e^{-2^\symX}
e^{-2\pi\symImaginary\symIndexL\symX}d\symX.
\end{multline}
The variable transformation $\symY = 2^\symX$ yields
\begin{equation}
\symA_\symIndexL
=
2\int_{0}^{\infty} e^{-\symY}
\symY^{-\frac{2\pi\symImaginary\symIndexL}{\log(2)}}d\symY
= 
2\Gamma\!\left(1-\frac{2\pi\symImaginary\symIndexL}{\log(2)}\right)
\end{equation}
where $\Gamma$ denotes the gamma function. Obviously, the constant term in the Fourier series is equal to $1$, because $\symA_0=2$.

In order to investigate the maximum deviation of $\symPowerSeriesFunc(\symX)$ from this value we consider all further coefficients $\symA_\symIndexL$ with $\symIndexL\geq 1$. Using the identity $\left|\Gamma(1+\symImaginary\symX)\right| = 
\sqrt{\frac{\pi\symX}{\sinh(\pi\symX)}}
$
we are able to write for the absolute values of the coefficients
\begin{equation}
|\symA_\symIndexL| = 
2\sqrt{\frac{\symB\symIndexL}{\sinh(\symB\symIndexL)}}
\quad
\text{with}
\
\symB:=\frac{2\pi^2}{\log 2}.
\end{equation}
In particular, the amplitude of the first harmonic is 
$|\symA_1|
\approx
\num{9.884e-6}$. Clearly, the maximum deviation of $\symPowerSeriesFunc(\symX)-1$ from the first harmonic must be smaller than $\sum_{\symIndexL=2}^{\infty} 
|\symA_\symIndexL|$. The ratio of subsequent coefficients is given by
\begin{equation}
\frac{|\symA_{\symIndexL+1}|}{|\symA_{\symIndexL}|}
=
\sqrt{\frac{\symIndexL+1}{\symIndexL}}
\sqrt{\frac{
\sinh(\symB \symIndexL)
}{
\sinh(\symB (\symIndexL+1))
}}
=
\sqrt{\frac{\symIndexL+1}{\symIndexL}}
\sqrt{\frac{
1
}{
\cosh(\symB)
+
\frac{
\sinh(\symB)
}
{
\tanh(\symB \symIndexL)
}
}}.
\end{equation}
For $\symIndexL\geq 2$ we have $\sqrt{\frac{\symIndexL+1}{\symIndexL}}\leq \sqrt{\frac{3}{2}}$. Together with $\tanh(x)\leq 1$ we obtain
\begin{equation}
\frac{|\symA_{\symIndexL+1}|}{
|\symA_{\symIndexL}|
}
\leq
\sqrt{\frac{3}{2}}
\sqrt{\frac{
1
}{
\cosh(\symB)
+
\sinh(\symB)
}}
\\
=
\sqrt{\frac{
3
}{
2 e^\symB
}}
\end{equation}
which leads to $|\symA_{\symIndexL}| \leq 
|\symA_{2}|
\left(\sqrt{\frac{
3
}{
2 e^\symB
}}\right)^{\symIndexL-2}$ for $\symIndexL\geq 2$ and further to
\begin{equation}
\sum_{\symIndexL=2}^{\infty} 
|\symA_\symIndexL|
\leq
|\symA_2|
\sum_{\symIndexL=0}^{\infty} 
\left(
\sqrt{\frac{
3
}{
2e^\symB
}}
\right)^\symIndexL
=
2\sqrt{
\frac{2\symB}{\sinh(2\symB)}}
\frac{1}{1-
\sqrt{\frac{
3
}{
2 e^\symB
}}
}
\approx
\num{9.154e-12}.
\end{equation}
As a consequence, the maximum deviation of $\symPowerSeriesFunc(\symX)$ from 1 is bounded by
\begin{equation}
\num{9.884e-6}
\leq
|\symA_1|
-
\sum_{\symIndexL=2}^{\infty} 
|\symA_\symIndexL|
\leq
\max_{\symX}(|\symPowerSeriesFunc(\symX)-1|) 
\leq 
|\symA_1|
+
\sum_{\symIndexL=2}^{\infty} 
|\symA_\symIndexL|
\leq 
\num{9.885e-6}.
\end{equation}
